---
title: "C4_O2_Behavior"
author: "Alan Ivory"
date: "`r Sys.Date()`"
output: html_document
---

# Set-up
Install necessary packages and import appropriate data

```{r setup_All}
pacman::p_load(tidyverse, readxl, raster,
               vegan, tigris, sf, sjPlot,
               sp, spOccupancy, ggrepel,
               lme4, lmerTest, MuMIn,
               brms, MCMCvis, cmdstanr,
               lubridate, forcats, tidybayes,
               ggdist, knitr, kableExtra,
               officer, flextable, purrr,
               posterior, patchwork, scales)

# Install dependencies
#install.packages(c("posterior", "RcppParallel", "jsonlite"))

# Install cmdstanr
#install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))


# Tree PCQ Data
tree_data <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                        sheet = "Tree_PCQ")

# Soil Data
fuel_data <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                        sheet = "Fuel_Sampling")

# Veg Data
Veg_Cover <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                        sheet = "Veg_Cover")

# Shrub Cover Data
shrub_data <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                         sheet = "Shrub_Cover")

# Site Data
CameraData <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/04_Wildlife/02_Data/CameraData.xlsx")

CameraLoc <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/04_Wildlife/02_Data/CameraLoc.xlsx",
                  sheet = "CameraLocations")

# Add effort data
effort_matrix <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/04_Wildlife/02_Data/CameraLoc.xlsx",
                            sheet = "Effort_Matrix_Full") %>%
  pivot_longer(cols = matches("^202[4-5]-"), names_to = "week", values_to = "days") %>%
  filter(days == "7") %>%
  dplyr::select(Plot, week)
```

## Number of quadrats sampled per plot
```{r Quadrats_All}
# Count the total number of quadrats per plot
quadrat_count <- Veg_Cover %>%
  group_by(Plot) %>%
  summarize(total_quadrats = n_distinct(Quadrat), .groups = "drop")
```

## Filter All data to only include specified species (Per PLANTS database)
```{r Filter_All}
#Filter tree data to only include trees with "tree" in the growth column
tree_data <- dplyr::filter(tree_data, Growth == "Tree")

#Filter Veg Cover to exclude Shrubs and Trees
Veg_Cover <- dplyr::filter(Veg_Cover, Growth != "Shrub" & Growth != "Tree")

#Filter Shrub Cover to only include Shrubs and Trees
shrub_data <- dplyr::filter(shrub_data, Growth == "Shrub" | Growth == "Tree")

```

## Place holder in case we need to filter vegetation by frequency
```{r RareFilter_All}
# Total number of sites
total_sites <- nrow(CameraLoc)

# Function to filter data by frequency
filter_by_frequency <- function(df) {
  # Group data by species and calculate the frequency
  freq <- df %>%
    group_by(Species) %>%
    summarise(Frequency = n_distinct(Plot) / nrow(CameraLoc) * 100) %>%
    filter(Frequency >= 0)
  
  # Filter the original data to include only species with frequency >= XX%
  filtered_df <- df %>%
    filter(Species %in% freq$Species)
  
  return(filtered_df)
}

# Filter tree data by frequency
tree_data <- filter_by_frequency(tree_data)

# Filter Veg Cover data by frequency
Veg_Cover <- filter_by_frequency(Veg_Cover)

# Filter Shrub Cover data by frequency
shrub_data <- filter_by_frequency(shrub_data)

```



## Shrub Cover Conversion
```{r Shrub_Cover_All}
# Total length of Shrub cover at a site
shrub_cover <- shrub_data %>%
  mutate(Cover = Line_End - Line_Start) %>%
  group_by(Species_Name, Plot) %>%
  summarise(Shrub_Total_Cover = sum(Cover, na.rm = TRUE), .groups = "drop") %>%
  mutate(Shrub_Percent_Cover = Shrub_Total_Cover / 3000 * 100)

# Summed length of shrub over at a site
shrub_cover_summed <- shrub_cover %>%
  group_by(Plot) %>%
  summarize(total_shrub_cover = sum(Shrub_Total_Cover, na.rm = TRUE), .groups = "drop")

```

## Herbacous Cover Conversion
```{r Cogon_Site_All}
# Combine Plot and Quadrat columns
Veg_Cover <- Veg_Cover %>%
  mutate(Plot_Quadrat = paste(Plot, Quadrat, sep = '_'))

# Join with CogonSites to get site information
Veg_Cover <- Veg_Cover %>%
  left_join(CameraLoc, by = "Plot")

# Sum species cover across quadrats for each species at each plot
veg_cover_summed <- Veg_Cover %>%
  group_by(Plot, Species_Name) %>%
  summarize(total_cover = sum(Cover_Per, na.rm = TRUE), .groups = "drop")

# Calculate average herbaceous species cover
avg_species_cover <- veg_cover_summed %>%
  left_join(quadrat_count, by = "Plot") %>%
  mutate(avg_cover = total_cover / total_quadrats)

```

## Merging Herb cover with Shrub
```{r Herb_Shrub_All}

# Merge shrub cover with herbaceous average cover
combined_cover <- avg_species_cover %>%
  full_join(
    shrub_cover %>%
      dplyr::select(Plot, Species_Name, Shrub_Percent_Cover),
    by = c("Plot", "Species_Name")
  ) %>%
  mutate(
    overlap_flag = ifelse(!is.na(avg_cover) & !is.na(Shrub_Percent_Cover), TRUE, FALSE), # Flag overlaps
    final_cover = case_when(
      !is.na(avg_cover) & is.na(Shrub_Percent_Cover) ~ avg_cover,  # Use herbaceous cover if no shrub data
      is.na(avg_cover) & !is.na(Shrub_Percent_Cover) ~ Shrub_Percent_Cover, # Use shrub cover if no herbaceous data
      TRUE ~ NA_real_ # Leave as NA where overlaps exist
    )
  )

# Species Matrix
species_matrix <- combined_cover %>%
  dplyr::select(Plot, Species_Name, final_cover) %>%
  pivot_wider(
    names_from = Species_Name,
    values_from = final_cover,
    values_fill = 0
  )
```

## Summarize Cogongrass Cover
```{r Cogon_Cover_All}
avg_cogongrass_cover <- species_matrix %>%
  group_by(Plot) %>%
  summarize(Avg_Cogongrass_Cover = sum(Imperata_cylindrica, na.rm = TRUE) / n(), .groups = "drop")

```

## Herbacous Shannon Diversity Index
```{r Shannon_Diversity}
# Summarize species cover by site
site_species_cover <- Veg_Cover %>%
  group_by(Plot, Species_Name) %>%
  summarize(total_cover = sum(Cover_Per, na.rm = TRUE)) %>%
  ungroup()

# Calculate Shannon diversity per site
Veg_shannon_diversity <- site_species_cover %>%
  group_by(Plot) %>%
  mutate(proportion = total_cover / sum(total_cover)) %>%
  summarize(Veg_shannon_index = -sum(proportion * log(proportion), na.rm = TRUE))

print(Veg_shannon_diversity)
```

## Vegetation Height
```{r Veg_Height_All}
if (!is.numeric(fuel_data$Height)) {
  fuel_data$Height <- as.numeric(as.character(fuel_data$Height))
}

# Calculate average vegetation height per plot
veg_height <- fuel_data %>%
  group_by(Plot) %>%
  summarize(avg_veg_height = mean(Height, na.rm = TRUE), .groups = "drop")
```

## Tree Metrics
```{r Tree_Metrics_All}
# Tree density from point-centered quarter data
if (!is.numeric(tree_data$Distance)) {
  tree_data$Distance <- as.numeric(as.character(tree_data$Distance))
}

tree_density_data <- tree_data %>%
  group_by(Plot) %>%
  summarize(Average_Distance = mean(Distance) / 100,  # Convert to meters
            Tree_Density = 10000 / (Average_Distance^2))  # Convert to trees per hectare

# Average canopy cover from vegetation quadrats
tree_canopy_data <- Veg_Cover %>%
  distinct(Plot, Quadrat, .keep_all = TRUE) %>%  # Ensure each quadrat counts once per plot
  group_by(Plot) %>%
  summarize(Avg_Canopy_Cover = mean(Canopy_Cover, na.rm = TRUE), .groups = "drop") # Calculate the average canopy cover per plot

cor(tree_density_data$Tree_Density, tree_canopy_data$Avg_Canopy_Cover)


```

# Objective 1: Assessing if cogongrass invasion-related factors influence wildlife presence

## Add Covariates
```{r Cov_Dat}
CameraLoc <- CameraLoc %>%
  left_join(Veg_shannon_diversity, by = "Plot") %>%
  left_join(avg_cogongrass_cover, by = "Plot") %>%
  left_join(shrub_cover_summed %>% dplyr::select(Plot, total_shrub_cover), by = "Plot") %>%
  left_join(veg_height, by = "Plot") %>%
  left_join(tree_density_data %>% dplyr::select(Plot, Tree_Density), by = "Plot") %>%
  left_join(tree_canopy_data %>% dplyr::select(Plot, Avg_Canopy_Cover), by = "Plot") %>%
  dplyr::select(-Authority)
```

## Subset to mammals and turkeys with greater than 10 observations
```{r Sub_Species}
# Group by Name and count the number of observations
species_counts <- CameraData %>%
  filter(Class == "Mammalia" | 
  Name == "Meleagris_gallopavo") %>%
  group_by(Name) %>%
  summarize(count = n(), .groups = "drop")

# Filter for species with count greater than 10
species_subset <- species_counts %>%
  filter(count > 10) %>%
  pull(Name)

# Filter CameraData to only include species with count > 10
CameraData <- CameraData %>%
  filter(Name %in% species_subset)
```

## Create Detection Matrix
```{r DetMat}
# Format Data Weekly
observations_weekly <- CameraData %>%
  group_by(Plot, week = format(as.Date(Date), "%Y-%U"), Name) %>%
  summarise(observations = n(), .groups = 'drop')

# Merge with Effort Matrix to include only valid weeks
observations_weekly <- effort_matrix %>%
  left_join(observations_weekly, by = c("Plot" = "Plot", "week")) %>%
  replace_na(list(observations = 0))

# Convert to wide format
observations_species <- observations_weekly %>%
  pivot_wider(names_from = Name, values_from = observations, values_fill = list(observations = 0)) %>%
  dplyr::select(-"NA")

# Create detection array
site_names <- sort(unique(observations_species$Plot))
species_names <- setdiff(colnames(observations_species), c("Plot", "week"))

num_sites <- length(site_names)
num_weeks <- length(unique(observations_species$week))
num_species <- length(species_names)

detection_array <- array(0, dim = c(num_sites, num_weeks, num_species))
dimnames(detection_array) <- list(site_names, unique(observations_species$week), species_names)

for (s in seq_along(species_names)) {
  species_col <- species_names[s]
  for (i in seq_along(site_names)) {
    site <- site_names[i]
    for (j in seq_along(unique(observations_species$week))) {
      week <- unique(observations_species$week)[j]
      detection_array[i, j, s] <- ifelse(
        any(observations_species$Plot == site & observations_species$week == week & observations_species[[species_col]] > 0),
        1, 0
      )
    }
  }
}

dim(detection_array) # Should be num_sites x num_weeks x num_species
```

## Prepare Site Covariates
```{r SiteCov}
# Duplicate CameraLoc to be used later
CameraLoc_O2 <- CameraLoc

# Standardize the covariates
CameraLoc <- CameraLoc %>%
  dplyr::select(-Plot, -Camera, -Lat, -Long, -Status, - Start_Date)
covariates_matrix <- as.matrix(CameraLoc)
rownames(covariates_matrix) <- site_names

# Standardizing covariates
covariates_matrix <- scale(covariates_matrix)
```

## Combine data into a list
This requires combining the presence data and the site covariate data into a single list. This also means that the presence data is in a 3-d format.
```{r Combine_List_Data}
# Combine into a named list
data_list <- list(
  y = detection_array,
  occ.covs = covariates_matrix,
  det.covs = covariates_matrix
)
```

## Convert occupancy and detection covariates to a dataframe (from matrix)
```{r Convert_DF}
# Convert occupancy and detection covariates to a dataframe
data_list[["occ.covs"]] <- as.data.frame(data_list[["occ.covs"]])
data_list[["occ.covs"]]$total_shrub_cover <- as.numeric(data_list[["occ.covs"]]$total_shrub_cover)

data_list[["det.covs"]] <- as.data.frame(data_list[["det.covs"]])
data_list[["det.covs"]]$total_shrub_cover <- as.numeric(data_list[["det.covs"]]$total_shrub_cover)

# Make species the first dimension
data_list$y <- aperm(data_list$y, c(3, 1, 2))

dimnames(data_list$y) <- list(species = dimnames(data_list$y)[[1]],
                              site = dimnames(data_list$y)[[2]],
                              week = dimnames(data_list$y)[[3]])

str(data_list)
```

# Objective 2: Determine whether wildlife behaviors differ in and around cogongrass patches, specifically as it relates to whether specific taxa avoid cogongrass invaded areas.

## Behavior Occurence
```{r Behavior_Occurence}
CameraData <- CameraData%>%
  dplyr::select(-Status)

O2_data <- CameraData %>%
  left_join(CameraLoc_O2, by = "Plot")

O2_data <- O2_data %>%
  mutate(
    DateTime = update(Date,
                      hour   = hour(Time),
                      minute = minute(Time),
                      second = second(Time))
  )

gap_mins <- 30

O2_data <- O2_data %>%
  filter(!is.na(DateTime)) %>%
  arrange(Plot, Name, DateTime) %>%
  group_by(Plot, Name) %>%
  group_modify(~{
    df <- .x
    keep <- logical(nrow(df))
    last_kept <- as.POSIXct(NA, tz = tz(df$DateTime[1]))
    for (i in seq_len(nrow(df))) {
      if (is.na(last_kept) || difftime(df$DateTime[i], last_kept, units = "mins") > gap_mins) {
        keep[i] <- TRUE
        last_kept <- df$DateTime[i]
      }
    }
    df[keep, , drop = FALSE]
  }) %>%
  ungroup()

dat <- O2_data %>%
  filter(!is.na(Behavior), !is.na(Status), !is.na(BehLoc)) %>%
  mutate(
    # time-of-day as proportion of a 24-hour day, then z-score and quadratic
    time_prop = (hour(DateTime)*3600 + minute(DateTime)*60 + second(DateTime)) / 86400,
    time_z = as.numeric(scale(time_prop)),
    time_z2 = time_z^2,
    # temperature standardized
    temp_z = as.numeric(scale(Air.TemperatureC)),
    # month factor to absorb temporal clustering
    Month = factor(month(DateTime)),
    # set factor baselines
    Behavior = factor(Behavior, ordered = FALSE),
    Behavior = relevel(Behavior, ref = "Local_Search"),
    Status   = factor(Status, levels = c("Non_Invaded","Invaded")),
    BehLoc   = factor(BehLoc, levels = c("Non_Patch","Patch")),  # Non_Patch vs Patch
    Plot     = factor(Plot),
    Name     = factor(Name),
    Site     = factor(Site),
    Camera.Type = factor(Camera.Type)
  )

```

```{r models}
priors <- c(
  prior(normal(0, 1),   class = "b",         dpar = "muForaging"),
  prior(normal(0, 1),   class = "b",         dpar = "muTransit"),
  prior(normal(0, 2.5), class = "Intercept", dpar = "muForaging"),
  prior(normal(0, 2.5), class = "Intercept", dpar = "muTransit"),
  prior(exponential(1), class = "sd",        dpar = "muForaging"),
  prior(exponential(1), class = "sd",        dpar = "muTransit")
)

# Objective 2a: cover effect on behavior across all sites
model_status <- brm(
  bf(Behavior ~ Avg_Cogongrass_Cover * Name +
       (1 | Site) + (1 | Camera.Type) + (1 | Month)),
  family  = categorical(link = "logit", refcat = "Local_Search"),
  data    = dat,
  prior   = priors,
  chains  = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12)
)

# Objective 2b: does the cover effect differ inside vs outside patches?
dat_inv <- dat %>% filter(Status == "Invaded")

model_loc <- brm(
  bf(Behavior ~ BehLoc * Avg_Cogongrass_Cover * Name +
       (1 | Site) + (1 | Camera.Type) + (1 | Month)),
  family  = categorical(link = "logit", refcat = "Local_Search"),
  data    = dat_inv,
  prior   = priors,
  chains  = 4, cores = 4, iter = 4000, warmup = 2000,
  control = list(adapt_delta = 0.95, max_treedepth = 12)
)

# Diagnostics
summary(model_status); summary(model_loc)
loo(model_status);     loo(model_loc)
```


# Helper Functions

```{r helpers}
# Probability of direction
pd_fn <- function(x) max(mean(x > 0), mean(x < 0))

# Display labels
common_names <- c(
  "Canis_latrans"          = "Coyote",
  "Odocoileus_virginianus" = "White-Tailed Deer",
  "Dasypus_novemcinctus"   = "Nine-Banded Armadillo",
  "Procyon_lotor"          = "Raccoon",
  "Lynx_rufus"             = "Bobcat",
  "Didelphis_virginiana"   = "Virginia Opossum",
  "Sylvilagus_floridanus"  = "Eastern Cottontail Rabbit",
  "Meleagris_gallopavo"    = "Wild Turkey",
  "Sciurus_carolinensis"   = "Gray Squirrel"
)

behavior_colors <- c("Local Search" = "#0072B2",
                     "Foraging"     = "#009E73",
                     "Transit"      = "#D55E00")
location_colors <- c("Non-Patch"    = "#0072B2",
                     "Patch"        = "#D55E00")


save_effect_docx <- function(df, file_name, caption_text) {
  if (".category" %in% names(df)) {
    df <- df %>%
      mutate(Behavior = factor(.category,
               levels = c("Local_Search", "Foraging", "Transit"),
               labels = c("Local Search", "Foraging", "Transit"))) %>%
      dplyr::select(-.category)
  }
  if ("Name" %in% names(df)) {
    df <- df %>%
      mutate(Species = recode(as.character(Name), !!!common_names)) %>%
      dplyr::select(-Name)
  }

  bold_flag <- df$pd >= 0.95

  out <- df %>%
    mutate(`Median (pp)`   = round(median_pp, 2),
           pd              = round(pd, 2),
           `Lower 95% CrI` = round(lower, 2),
           `Upper 95% CrI` = round(upper, 2)) %>%
    dplyr::select(any_of(c("Species", "BehLoc", "Behavior")),
                  `Median (pp)`, pd, `Lower 95% CrI`, `Upper 95% CrI`)

  ft <- flextable(out) %>%
    set_caption(caption_text) %>%
    bold(i = which(bold_flag), bold = TRUE) %>%   # bold credible rows
    autofit()
  save_as_docx(ft, path = file_name)
  list(table = out, flextable = ft)
}
```


# 3. Parameter tables  (APPENDIX — conditional on reference species/behavior)

```{r parameter_tables}

make_pd_table <- function(model, file_name, caption_text) {
  draws <- as_draws_df(model)
  beta  <- draws[, grep("^b_", colnames(draws)), drop = FALSE]


  tab <- tibble::tibble(
    parameter = gsub("^b_", "", colnames(beta)),
    mean      = colMeans(beta),
    pd        = apply(beta, 2, pd_fn),
    lower     = apply(beta, 2, quantile, probs = 0.025),
    upper     = apply(beta, 2, quantile, probs = 0.975)
  ) %>%
    mutate(pd_95 = pd >= 0.95)

  disp <- tab %>%
    transmute(Parameter       = parameter,
              Mean            = round(mean, 2),
              pd              = round(pd, 2),
              `Lower 95% CrI` = round(lower, 2),
              `Upper 95% CrI` = round(upper, 2))

  ft <- flextable(disp) %>%
    set_caption(caption_text) %>%
    bold(i = which(tab$pd_95), bold = TRUE) %>%
    autofit()
  save_as_docx(ft, path = file_name)
  list(table = tab, flextable = ft)
}

pd_status <- make_pd_table(model_status, "PD_Table_Status_Model.docx",
                           "Probability of Direction - Status Model")
pd_loc    <- make_pd_table(model_loc,    "PD_Table_Location_Model.docx",
                           "Probability of Direction - Location Model")


make_pub_table_with_ci <- function(model, pd_table, file_name, caption_text) {
  ci_df <- as.data.frame(summary(model)$fixed) %>%
    tibble::rownames_to_column("parameter") %>%
    dplyr::select(parameter, Estimate, `l-95% CI`, `u-95% CI`) %>%
    rename(mean = Estimate, l_95 = `l-95% CI`, u_95 = `u-95% CI`)

  df_pub <- pd_table %>%
    dplyr::select(parameter, pd) %>%
    left_join(ci_df, by = "parameter") %>%
    mutate(bold_flag       = pd >= 0.95,           
           Mean            = round(mean, 2),
           pd              = round(pd, 2),
           `Lower 95% CrI` = round(l_95, 2),
           `Upper 95% CrI` = round(u_95, 2)) %>%
    dplyr::select(Parameter = parameter, Mean, pd,
                  `Lower 95% CrI`, `Upper 95% CrI`, bold_flag)

  ft <- flextable(df_pub %>% dplyr::select(-bold_flag)) %>%
    set_caption(caption_text) %>%
    bold(i = which(df_pub$bold_flag), j = 1:5, bold = TRUE) %>%
    autofit()
  save_as_docx(ft, path = file_name)
  ft
}

ft_status <- make_pub_table_with_ci(model_status, pd_status$table,
              "Status_Model_Parameters_CI.docx",
              "Status Model Parameter Estimates with 95% Credible Intervals")
ft_loc    <- make_pub_table_with_ci(model_loc, pd_loc$table,
              "Location_Model_Parameters_CI.docx",
              "Location Model Parameter Estimates with 95% Credible Intervals")
```


# 4. Effect tables

```{r effect_tables_status}
cover_lo <- min(dat$Avg_Cogongrass_Cover, na.rm = TRUE)
cover_hi <- max(dat$Avg_Cogongrass_Cover, na.rm = TRUE)

newdata_effect <- expand_grid(
  cover_level = c("low", "high"),
  Name        = unique(dat$Name)
) %>%
  mutate(Avg_Cogongrass_Cover = ifelse(cover_level == "low", cover_lo, cover_hi),
         Site = NA, Camera.Type = NA, Month = NA)

eff_draws <- add_epred_draws(newdata_effect, model_status, re_formula = NA) %>%
  ungroup()

# Community-level
community_effect <- eff_draws %>%
  group_by(.draw, .category, cover_level) %>%
  summarise(.epred = mean(.epred), .groups = "drop") %>%
  pivot_wider(names_from = cover_level, values_from = .epred) %>%
  mutate(delta = high - low) %>%
  group_by(.category) %>%
  summarise(median_pp = median(delta) * 100,
            lower = quantile(delta, .025) * 100,
            upper = quantile(delta, .975) * 100,
            pd    = pd_fn(delta), .groups = "drop")

# Per species
species_effect <- eff_draws %>%
  pivot_wider(id_cols = c(.draw, .category, Name),
              names_from = cover_level, values_from = .epred) %>%
  mutate(delta = high - low) %>%
  group_by(Name, .category) %>%
  summarise(median_pp = median(delta) * 100,
            lower = quantile(delta, .025) * 100,
            upper = quantile(delta, .975) * 100,
            pd    = pd_fn(delta), .groups = "drop")

save_effect_docx(community_effect, "Effect_Community_Status.docx",
  "Community-level change in behavior probability across the cover gradient (Obj. 2a)")
save_effect_docx(species_effect,   "Effect_Species_Status.docx",
  "Species-specific change in behavior probability across the cover gradient (Obj. 2a)")
```

```{r effect_tables_location}
cover_lo_i <- min(dat_inv$Avg_Cogongrass_Cover, na.rm = TRUE)
cover_hi_i <- max(dat_inv$Avg_Cogongrass_Cover, na.rm = TRUE)

newdata_loc_eff <- expand_grid(
  BehLoc      = c("Non_Patch", "Patch"),
  cover_level = c("low", "high"),
  Name        = unique(dat_inv$Name)
) %>%
  mutate(Avg_Cogongrass_Cover = ifelse(cover_level == "low", cover_lo_i, cover_hi_i),
         Site = NA, Camera.Type = NA, Month = NA)

loc_eff_draws <- add_epred_draws(newdata_loc_eff, model_loc, re_formula = NA) %>%
  ungroup()

# --- Community (averaged over species) ---
loc_draws <- loc_eff_draws %>%
  group_by(.draw, .category, BehLoc, cover_level) %>%
  summarise(.epred = mean(.epred), .groups = "drop") %>%
  pivot_wider(names_from = cover_level, values_from = .epred) %>%
  mutate(cover_effect = high - low)

# cover effect within each location
loc_effect <- loc_draws %>%
  group_by(.category, BehLoc) %>%
  summarise(median_pp = median(cover_effect) * 100,
            lower = quantile(cover_effect, .025) * 100,
            upper = quantile(cover_effect, .975) * 100,
            pd    = pd_fn(cover_effect), .groups = "drop")

# Patch - Non_Patch
loc_diff <- loc_draws %>%
  dplyr::select(.draw, .category, BehLoc, cover_effect) %>%
  pivot_wider(names_from = BehLoc, values_from = cover_effect) %>%
  mutate(diff = Patch - Non_Patch) %>%
  group_by(.category) %>%
  summarise(median_pp = median(diff) * 100,
            lower = quantile(diff, .025) * 100,
            upper = quantile(diff, .975) * 100,
            pd    = pd_fn(diff), .groups = "drop")

# --- Per species ---
loc_draws_sp <- loc_eff_draws %>%
  pivot_wider(id_cols = c(.draw, .category, Name, BehLoc),
              names_from = cover_level, values_from = .epred) %>%
  mutate(cover_effect = high - low)

loc_effect_sp <- loc_draws_sp %>%
  group_by(Name, .category, BehLoc) %>%
  summarise(median_pp = median(cover_effect) * 100,
            lower = quantile(cover_effect, .025) * 100,
            upper = quantile(cover_effect, .975) * 100,
            pd    = pd_fn(cover_effect), .groups = "drop")

loc_diff_sp <- loc_draws_sp %>%
  dplyr::select(.draw, .category, Name, BehLoc, cover_effect) %>%
  pivot_wider(names_from = BehLoc, values_from = cover_effect) %>%
  mutate(diff = Patch - Non_Patch) %>%
  group_by(Name, .category) %>%
  summarise(median_pp = median(diff) * 100,
            lower = quantile(diff, .025) * 100,
            upper = quantile(diff, .975) * 100,
            pd    = pd_fn(diff), .groups = "drop")


# --- Community-level Patch -------------------
loc_level_draws <- loc_eff_draws %>%
  group_by(.draw, .category, BehLoc) %>%
  summarise(.epred = mean(.epred), .groups = "drop") %>%
  pivot_wider(names_from = BehLoc, values_from = .epred) %>%
  mutate(level_diff = Patch - Non_Patch)

loc_level <- loc_level_draws %>%
  group_by(.category) %>%
  summarise(median_pp = median(level_diff) * 100,
            lower = quantile(level_diff, .025) * 100,
            upper = quantile(level_diff, .975) * 100,
            pd    = pd_fn(level_diff), .groups = "drop")

# --- Species-specific version --------------------------------------------
loc_level_sp_draws <- loc_eff_draws %>%
  group_by(.draw, .category, Name, BehLoc) %>%
  summarise(.epred = mean(.epred), .groups = "drop") %>%   # avg over cover_level
  pivot_wider(names_from = BehLoc, values_from = .epred) %>%
  mutate(level_diff = Patch - Non_Patch)

loc_level_sp <- loc_level_sp_draws %>%
  group_by(Name, .category) %>%
  summarise(median_pp = median(level_diff) * 100,
            lower = quantile(level_diff, .025) * 100,
            upper = quantile(level_diff, .975) * 100,
            pd    = pd_fn(level_diff), .groups = "drop")

save_effect_docx(loc_effect,    "Effect_Location_byLocation.docx",
  "Cover effect on behavior within each location (Obj. 2b)")
save_effect_docx(loc_diff,      "Effect_Location_PatchDiff.docx",
  "Patch minus Non-Patch difference in the cover effect (Obj. 2b headline)")
save_effect_docx(loc_effect_sp, "Effect_Location_Species_byLocation.docx",
  "Species-specific cover effect within each location (Obj. 2b)")
save_effect_docx(loc_diff_sp,   "Effect_Location_Species_PatchDiff.docx",
  "Species-specific Patch minus Non-Patch difference in the cover effect (Obj. 2b)")

save_effect_docx(loc_level,    "Effect_Location_LevelDiff.docx",
  "Patch minus Non-Patch difference in behavior probability (Obj. 2b, level contrast)")
save_effect_docx(loc_level_sp, "Effect_Location_Species_LevelDiff.docx",
  "Species-specific Patch minus Non-Patch difference in behavior probability (Obj. 2b)")
```


# 5. Figures

```{r cover_gradient_community}
cover_seq <- seq(cover_lo, cover_hi, length.out = 100)

behavior_summary <- add_epred_draws(
    expand_grid(Avg_Cogongrass_Cover = cover_seq, Name = unique(dat$Name)) %>%
      mutate(Site = NA, Camera.Type = NA, Month = NA),
    model_status, re_formula = NA, ndraws = 2000) %>%
  ungroup() %>%
  group_by(Avg_Cogongrass_Cover, .draw, .category) %>%
  summarise(.epred = mean(.epred), .groups = "drop") %>%
  group_by(Avg_Cogongrass_Cover, .category) %>%
  median_qi(.epred, .width = 0.95) %>%
  mutate(Behavior = factor(.category,
           levels = c("Local_Search", "Foraging", "Transit"),
           labels = c("Local Search", "Foraging", "Transit")))

plot_proportions <- dat %>%
  group_by(Plot) %>%
  mutate(plot_cover = mean(Avg_Cogongrass_Cover, na.rm = TRUE)) %>%
  group_by(Plot, plot_cover, Behavior) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Plot, plot_cover) %>%
  mutate(site_prob = n / sum(n)) %>%
  ungroup() %>%
  mutate(Behavior = factor(Behavior,
           levels = c("Local_Search", "Foraging", "Transit"),
           labels = c("Local Search", "Foraging", "Transit")))

p_community <- ggplot(behavior_summary,
       aes(Avg_Cogongrass_Cover, .epred, color = Behavior, fill = Behavior)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), alpha = 0.2, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(data = plot_proportions, inherit.aes = FALSE,
             aes(plot_cover, site_prob, fill = Behavior),
             shape = 21, color = "black", alpha = 0.3, size = 2.2, stroke = 0.6) +
  scale_color_manual(values = behavior_colors, name = "Behavior") +
  scale_fill_manual(values = behavior_colors, name = "Behavior") +
  scale_y_continuous(limits = c(0, 1), labels = label_percent(accuracy = 1)) +
  labs(x = "Cogongrass Cover (%)", y = "Predicted Probability") +
  theme_classic(base_size = 12) + theme(axis.title = element_text(face = "bold"))

print(p_community)
ggsave("Figure_Behavior_by_Cover.png", p_community,
       width = 7, height = 5, dpi = 600, bg = "white")
```

```{r cover_gradient_species}
behavior_summary_sp <- add_epred_draws(
    expand_grid(Avg_Cogongrass_Cover = cover_seq, Name = unique(dat$Name)) %>%
      mutate(Site = NA, Camera.Type = NA, Month = NA),
    model_status, re_formula = NA, ndraws = 2000) %>%
  ungroup() %>%
  group_by(Name, Avg_Cogongrass_Cover, .category) %>%
  median_qi(.epred, .width = 0.95) %>%
  mutate(Behavior = factor(.category,
           levels = c("Local_Search", "Foraging", "Transit"),
           labels = c("Local Search", "Foraging", "Transit")),
         Name = factor(recode(Name, !!!common_names), levels = common_names))

p_species <- ggplot(behavior_summary_sp,
       aes(Avg_Cogongrass_Cover, .epred, color = Behavior, fill = Behavior)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), alpha = 0.2, color = NA) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ Name) +
  scale_color_manual(values = behavior_colors, name = "Behavior") +
  scale_fill_manual(values = behavior_colors, name = "Behavior") +
  scale_y_continuous(limits = c(0, 1), labels = label_percent()) +
  labs(x = "Cogongrass Cover (%)", y = "Behavior Probability") +
  theme_classic()

print(p_species)
ggsave("Figure_Behavior_by_Species.png", p_species,
       width = 8, height = 7, dpi = 600, bg = "white")
```

```{r cover_gradient_location}
cover_seq_inv <- seq(cover_lo_i, cover_hi_i, length.out = 100)

behavior_summary_loc <- add_epred_draws(
    expand_grid(BehLoc = c("Non_Patch", "Patch"),
                Avg_Cogongrass_Cover = cover_seq_inv,
                Name = unique(dat_inv$Name)) %>%
      mutate(Site = NA, Camera.Type = NA, Month = NA),
    model_loc, re_formula = NA, ndraws = 2000) %>%
  ungroup() %>%
  group_by(BehLoc, Avg_Cogongrass_Cover, .draw, .category) %>%
  summarise(.epred = mean(.epred), .groups = "drop") %>%
  group_by(BehLoc, Avg_Cogongrass_Cover, .category) %>%
  median_qi(.epred, .width = 0.95) %>%
  mutate(Behavior = factor(.category,
           levels = c("Local_Search", "Foraging", "Transit"),
           labels = c("Local Search", "Foraging", "Transit")),
         BehLoc = factor(BehLoc, levels = c("Non_Patch", "Patch"),
                         labels = c("Non-Patch", "Patch")))

plot_proportions_loc <- dat_inv %>%
  group_by(Plot) %>%
  mutate(plot_cover = mean(Avg_Cogongrass_Cover, na.rm = TRUE)) %>%
  group_by(Plot, plot_cover, BehLoc, Behavior) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Plot, plot_cover, BehLoc) %>%
  mutate(site_prob = n / sum(n)) %>%
  ungroup() %>%
  mutate(BehLoc = factor(BehLoc, levels = c("Non_Patch", "Patch"),
                         labels = c("Non-Patch", "Patch")),
         Behavior = factor(Behavior,
           levels = c("Local_Search", "Foraging", "Transit"),
           labels = c("Local Search", "Foraging", "Transit")))

p_loc <- ggplot(behavior_summary_loc,
       aes(Avg_Cogongrass_Cover, .epred, color = BehLoc, fill = BehLoc)) +
  geom_ribbon(aes(ymin = .lower, ymax = .upper), alpha = 0.2, color = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(data = plot_proportions_loc, inherit.aes = FALSE,
             aes(plot_cover, site_prob, fill = BehLoc),
             shape = 21, color = "black", alpha = 0.3, size = 2.2, stroke = 0.6) +
  facet_wrap(~ Behavior) +
  scale_color_manual(values = location_colors, name = "Location") +
  scale_fill_manual(values = location_colors, name = "Location") +
  scale_y_continuous(limits = c(0, 1), labels = label_percent(accuracy = 1)) +
  labs(x = "Cogongrass Cover (%)", y = "Predicted Probability") +
  theme_classic(base_size = 12) + theme(axis.title = element_text(face = "bold"))

print(p_loc)
ggsave("Figure_Behavior_by_Location.png", p_loc,
       width = 8, height = 4, dpi = 600, bg = "white")
```

```{r cover_gradient_location_species}
behavior_summary_loc_sp <- add_epred_draws(
    expand_grid(BehLoc = c("Non_Patch", "Patch"),
                Avg_Cogongrass_Cover = cover_seq_inv,
                Name = unique(dat_inv$Name)) %>%
      mutate(Site = NA, Camera.Type = NA, Month = NA),
    model_loc, re_formula = NA, ndraws = 2000) %>%
  ungroup() %>%
  group_by(Name, BehLoc, Avg_Cogongrass_Cover, .category) %>%
  median_qi(.epred, .width = 0.95) %>%
  mutate(Behavior = factor(.category,
           levels = c("Local_Search", "Foraging", "Transit"),
           labels = c("Search", "Foraging", "Transit")),
         BehLoc = factor(BehLoc, levels = c("Non_Patch", "Patch"),
                         labels = c("Non-Patch", "Patch")),
         Name = factor(recode(Name, !!!common_names), levels = common_names))

make_loc_species_plot <- function(df) {
  ggplot(df, aes(Avg_Cogongrass_Cover, .epred, color = BehLoc, fill = BehLoc)) +
    geom_ribbon(aes(ymin = .lower, ymax = .upper), alpha = 0.2, color = NA) +
    geom_line(linewidth = 0.7) +
    facet_grid(Name ~ Behavior) +
    scale_color_manual(values = location_colors, name = "Location") +
    scale_fill_manual(values = location_colors, name = "Location") +
    scale_y_continuous(limits = c(0, 1), labels = label_percent()) +
    labs(x = "Cogongrass Cover (%)", y = "Behavior Probability") +
    theme_classic(base_size = 11)
}

species_levels <- levels(behavior_summary_loc_sp$Name)
p_loc_sp_1 <- make_loc_species_plot(behavior_summary_loc_sp %>%
                filter(Name %in% species_levels[1:5]))
p_loc_sp_2 <- make_loc_species_plot(behavior_summary_loc_sp %>%
                filter(Name %in% species_levels[6:length(species_levels)]))

print(p_loc_sp_1); print(p_loc_sp_2)
ggsave("Figure_Behavior_by_Species_Within_1.png", p_loc_sp_1,
       width = 9, height = 8, dpi = 600, bg = "white")
ggsave("Figure_Behavior_by_Species_Within_2.png", p_loc_sp_2,
       width = 9, height = 7, dpi = 600, bg = "white")
```

## Patch vs. Non-Patch Differences
```{r patch_nonpatch_diff}
library(forcats)

loc_level_draws <- loc_level_draws %>%
  mutate(Behavior = factor(.category,
           levels = c("Local_Search","Foraging","Transit"),
           labels = c("Local Search","Foraging","Transit")),
         level_diff_pp = level_diff * 100)

loc_level <- loc_level %>%
  mutate(Behavior = factor(.category,
           levels = c("Local_Search","Foraging","Transit"),
           labels = c("Local Search","Foraging","Transit")))

p_level <- ggplot(loc_level_draws,
       aes(x = level_diff_pp, y = fct_rev(Behavior), fill = Behavior)) +
  geom_vline(xintercept = 0, linetype = 2, color = "grey40") +
  stat_halfeye(.width = 0.95, point_interval = "median_qi",
               slab_alpha = 0.75, normalize = "xy") +
  geom_text(data = loc_level,
            aes(x = upper, y = fct_rev(Behavior),
                label = sprintf("pd = %.2f", pd)),
            inherit.aes = FALSE, hjust = -0.15, size = 3.3) +
  scale_fill_manual(values = behavior_colors, guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.18))) +
  labs(x = "Patch − Non-Patch Difference in Predicted Probability",
       y = NULL) +
  theme_classic(base_size = 12) +
  theme(axis.title = element_text(face = "bold"))

print(p_level)
ggsave("Figure_Patch_Level_Contrast.png", p_level,
       width = 7, height = 4.5, dpi = 600, bg = "white")
```

```{r combined_figure, eval=FALSE}
combined_plot <- (p_community / p_loc) + plot_annotation(tag_levels = "A")
ggsave("Figure_Behavior_Combined.png", combined_plot,
       width = 8, height = 10, dpi = 600, bg = "white")
```
