---
title: "Non-Spatial Occupancy Model"
author: "Alan Ivory"
date: "`r Sys.Date()`"
output: html_document
---

# Set-up
Install necessary packages and import appropriate data

```{r setup_All}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, readxl, raster, vegan, tigris,
               sf, sjPlot, sp, spOccupancy, ggrepel,
               lme4, lmerTest, MuMIn, brms, MCMCvis,
               bayesplot, patchwork, kableExtra,
               flextable, officer)

set.seed(97)

# Tree PCQ Data
tree_data <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                        sheet = "Tree_PCQ")

# CWD Data
cwd_data <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                       sheet = "CWD")

# Soil Data
fuel_data <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                        sheet = "Fuel_Sampling")

# Veg Data
Veg_Cover <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                        sheet = "Veg_Cover")

# Shrub Cover Data
shrub_data <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/05_SharedData/Field_Data_FL_AL_MS.xlsx",
                         sheet = "Shrub_Cover")

# Site Data
CameraData <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/04_Wildlife/02_Data/CameraData.xlsx")

CameraLoc <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/04_Wildlife/02_Data/CameraLoc.xlsx",
                  sheet = "CameraLocations")

# Add effort data
effort_matrix <- read_excel("C:/Users/DrewIvory/OneDrive - University of Florida/Desktop/School/PHD/01_Projects/04_Wildlife/02_Data/CameraLoc.xlsx",
                            sheet = "Effort_Matrix_Full") %>%
  pivot_longer(cols = matches("^202[4-5]-"), names_to = "week", values_to = "days") %>%
  filter(days == "7") %>%
  dplyr::select(Plot, week)
```


## Coordinate Transformation
```{r Coord_Transform_All}

coords_sf <- CameraLoc %>%
  dplyr::select(Plot, Long, Lat) %>%
  st_as_sf(coords = c("Long", "Lat"), crs = 4326) %>%
  st_transform(crs = 32616)  # UTM zone 16N

coords <- st_coordinates(coords_sf)
rownames(coords) <- CameraLoc$Plot
```

## Number of quadrats sampled per plot
```{r Quadrats_All}
# Count the total number of quadrats per plot
quadrat_count <- Veg_Cover %>%
  group_by(Plot) %>%
  summarize(total_quadrats = n_distinct(Quadrat), .groups = "drop")
```

## Filter All data to only include specified species (Per PLANTS database)
```{r Filter_All}
#Filter tree data to only include trees with "tree" in the growth column
tree_data <- dplyr::filter(tree_data, Growth == "Tree")

#Filter Veg Cover to exclude Shrubs and Trees
Veg_Cover <- dplyr::filter(Veg_Cover, Growth != "Shrub" & Growth != "Tree")

#Filter Shrub Cover to only include Shrubs and Trees
shrub_data <- dplyr::filter(shrub_data, Growth == "Shrub" | Growth == "Tree")

```

## Provides the option to rarify vegetative species
```{r RareFilter_All}
# Calculate the total number of sites
total_sites <- nrow(CameraLoc)

# Function to filter data by frequency
filter_by_frequency <- function(df) {
  # Group data by species and calculate the frequency
  freq <- df %>%
    group_by(Species) %>%
    summarise(Frequency = n_distinct(Plot) / nrow(CameraLoc) * 100) %>%
    filter(Frequency >= 0)
  
  # Filter the original data to include only species with frequency >= XX%
  filtered_df <- df %>%
    filter(Species %in% freq$Species)
  
  return(filtered_df)
}

# Filter tree data by frequency
tree_data <- filter_by_frequency(tree_data)

# Filter Veg Cover data by frequency
Veg_Cover <- filter_by_frequency(Veg_Cover)

# Filter Shrub Cover data by frequency
shrub_data <- filter_by_frequency(shrub_data)

```



## Shrub Cover Conversion
```{r Shrub_Cover_All}
# Total length of Shrub cover at a site
shrub_cover <- shrub_data %>%
  mutate(Cover = Line_End - Line_Start) %>%
  group_by(Species_Name, Plot) %>%
  summarise(Shrub_Total_Cover = sum(Cover, na.rm = TRUE), .groups = "drop") %>%
  mutate(Shrub_Percent_Cover = Shrub_Total_Cover / 3000 * 100)

# Summed length of shrub over at a site
shrub_cover_summed <- shrub_cover %>%
  group_by(Plot) %>%
  summarize(total_shrub_cover = sum(Shrub_Total_Cover, na.rm = TRUE), .groups = "drop")

```

## Herbacous Cover Conversion
```{r Cogon_Site_All}
# Combine Plot and Quadrat columns
Veg_Cover <- Veg_Cover %>%
  mutate(Plot_Quadrat = paste(Plot, Quadrat, sep = '_'))

# Join with CogonSites to get site information
Veg_Cover <- Veg_Cover %>%
  left_join(CameraLoc, by = "Plot")

# Sum species cover across quadrats for each species at each plot
veg_cover_summed <- Veg_Cover %>%
  group_by(Plot, Species_Name) %>%
  summarize(total_cover = sum(Cover_Per, na.rm = TRUE), .groups = "drop")

# Calculate average herbaceous species cover
avg_species_cover <- veg_cover_summed %>%
  left_join(quadrat_count, by = "Plot") %>%
  mutate(avg_cover = total_cover / total_quadrats)

```

## Merging Herb cover with Shrub

This species matrix includes herbaceous and shrub species
```{r Herb_Shrub_All}
# Merge shrub cover with herbaceous average cover
combined_cover <- avg_species_cover %>%
  full_join(
    shrub_cover %>%
      dplyr::select(Plot, Species_Name, Shrub_Percent_Cover),
    by = c("Plot", "Species_Name")
  ) %>%
  mutate(
    overlap_flag = ifelse(!is.na(avg_cover) & !is.na(Shrub_Percent_Cover), TRUE, FALSE), # Flag overlaps
    final_cover = case_when(
      !is.na(avg_cover) & is.na(Shrub_Percent_Cover) ~ avg_cover,  # Use herbaceous cover if no shrub data
      is.na(avg_cover) & !is.na(Shrub_Percent_Cover) ~ Shrub_Percent_Cover, # Use shrub cover if no herbaceous data
      TRUE ~ NA_real_ # NA where overlaps exist
    )
  )

# Species Matrix
species_matrix <- combined_cover %>%
  dplyr::select(Plot, Species_Name, final_cover) %>%
  pivot_wider(
    names_from = Species_Name,
    values_from = final_cover,
    values_fill = 0
  )
```

## Summarize Cogongrass Cover
```{r Cogon_Cover_All}
avg_cogongrass_cover <- species_matrix %>%
  group_by(Plot) %>%
  summarize(Avg_Cogongrass_Cover = sum(Imperata_cylindrica, na.rm = TRUE) / n(), .groups = "drop")

```

## Herbacous Shannon Diversity Index
```{r Shannon_Diversity}
# Summarize species cover by site
site_species_cover <- Veg_Cover %>%
  group_by(Plot, Species_Name) %>%
  summarize(total_cover = sum(Cover_Per, na.rm = TRUE)) %>%
  ungroup()

## Remove all Imperata_cylindrica_Live and Imperata_cylindrica from species
site_species_cover <- site_species_cover %>%
  filter(Species_Name != "Imperata_cylindrica_Live" & Species_Name != "Imperata_cylindrica")

# Calculate Shannon diversity per site
Veg_shannon_diversity <- site_species_cover %>%
  group_by(Plot) %>%
  mutate(proportion = total_cover / sum(total_cover)) %>%
  summarize(Veg_shannon_index = -sum(proportion * log(proportion), na.rm = TRUE))

print(Veg_shannon_diversity)
```

## Vegetation Height
```{r Veg_Height_All}
if (!is.numeric(fuel_data$Height)) {
  fuel_data$Height <- as.numeric(as.character(fuel_data$Height))
}

# Calculate average vegetation height per plot
veg_height <- fuel_data %>%
  group_by(Plot) %>%
  summarize(avg_veg_height = mean(Height, na.rm = TRUE), .groups = "drop")
```

## Tree Metrics
```{r Tree_Metrics_All}
# Tree density from point-centered quarter data
if (!is.numeric(tree_data$Distance)) {
  tree_data$Distance <- as.numeric(as.character(tree_data$Distance))
}

tree_density_data <- tree_data %>%
  group_by(Plot) %>%
  summarize(Average_Distance = mean(Distance) / 100,  # Convert to meters
            Tree_Density = 10000 / (Average_Distance^2))  # Convert to trees per hectare

# Average canopy cover from vegetation quadrats
tree_canopy_data <- Veg_Cover %>%
  distinct(Plot, Quadrat, .keep_all = TRUE) %>%  # Ensure each quadrat counts once per plot
  group_by(Plot) %>%
  summarize(Avg_Canopy_Cover = mean(Canopy_Cover, na.rm = TRUE), .groups = "drop") # Calculate the average canopy cover per plot

cor(tree_density_data$Tree_Density, tree_canopy_data$Avg_Canopy_Cover)


```

## CWD Clean
```{r CWD_Clean_All}
cwd_pa <- cwd_data %>%
  group_by(Plot) %>%
  summarise(
    CWD_presence = as.integer(any(Line_Start > 0 & Line_End > 0)),
    .groups = "drop"
  )
```

# Objective 1: Assessing if cogongrass invasion-related factors influence wildlife presence

## Add Covariates
```{r Cov_Dat}
CameraLoc <- CameraLoc %>%
  left_join(cwd_pa, by = "Plot") %>%
  left_join(Veg_shannon_diversity, by = "Plot") %>%
  left_join(avg_cogongrass_cover, by = "Plot") %>%
  left_join(shrub_cover_summed %>% dplyr::select(Plot, total_shrub_cover), by = "Plot") %>%
  left_join(veg_height, by = "Plot") %>%
  left_join(tree_density_data %>% dplyr::select(Plot, Tree_Density), by = "Plot") %>%
  left_join(tree_canopy_data %>% dplyr::select(Plot, Avg_Canopy_Cover), by = "Plot") %>%
  dplyr::select(-Authority)
```

## Distribution of Covariates
```{r Cov_Distribution}
par(mfrow = c(3, 3))
hist(CameraLoc$Cogon_Patch_Size, main = "Cogon Patch Size", xlab = "Size (m²)")
hist(CameraLoc$Veg_shannon_index, main = "Vegetation Shannon Index", xlab = "Shannon Index")
hist(CameraLoc$total_shrub_cover, main = "Total Shrub Cover", xlab = "Shrub Cover (%)")
hist(CameraLoc$Avg_Cogongrass_Cover, main = "Average Cogongrass Cover", xlab = "Cogongrass Cover (%)")
  
hist(CameraLoc$Tree_Density, main = "Tree Density", xlab = "Trees per ha")
hist(CameraLoc$Avg_Canopy_Cover, main = "Average Canopy Cover", xlab = "Canopy Cover (%)")
hist(CameraLoc$avg_veg_height, main = "Average Vegetation Height", xlab = "Height (cm)")
table(CameraLoc$CWD_presence)
par(mfrow = c(1, 1))
```

## Subset to mammals with >10 observations
```{r Sub_Species}
exclude_species <- c(
  "Odocoileus_virginianus",
  "Dasypus_novemcinctus"
)

species_counts <- CameraData %>%
  filter(
    (Class == "Mammalia" & !Name %in% exclude_species) |
    Name == "Meleagris_gallopavo"
  ) %>%
  group_by(Name) %>%
  summarize(count = n(), .groups = "drop")

# Filter for species with count greater than 10
species_subset <- species_counts %>%
  filter(count > 10) %>%
  pull(Name)

# Filter CameraData to only include species with count > 10
CameraData <- CameraData %>%
  filter(Name %in% species_subset)
```

## Create Detection Matrix
```{r DetMat}
# Format Data Weekly
observations_weekly <- CameraData %>%
  group_by(Plot, week = format(as.Date(Date), "%Y-%U"), Name) %>%
  summarise(observations = n(), .groups = 'drop')

# Merge with Effort Matrix to include only valid weeks
observations_weekly <- effort_matrix %>%
  left_join(observations_weekly, by = c("Plot" = "Plot", "week")) %>%
  replace_na(list(observations = 0))

# Convert to wide format
observations_species <- observations_weekly %>%
  pivot_wider(names_from = Name, values_from = observations, values_fill = list(observations = 0)) %>%
  dplyr::select(-"NA")

# Create detection array
site_names <- sort(unique(observations_species$Plot))
species_names <- setdiff(colnames(observations_species), c("Plot", "week"))

num_sites <- length(site_names)
num_weeks <- length(unique(observations_species$week))
num_species <- length(species_names)

detection_array <- array(0, dim = c(num_sites, num_weeks, num_species))
dimnames(detection_array) <- list(site_names, unique(observations_species$week), species_names)

for (s in seq_along(species_names)) {
  species_col <- species_names[s]
  for (i in seq_along(site_names)) {
    site <- site_names[i]
    for (j in seq_along(unique(observations_species$week))) {
      week <- unique(observations_species$week)[j]
      detection_array[i, j, s] <- ifelse(
        any(observations_species$Plot == site & observations_species$week == week & observations_species[[species_col]] > 0),
        1, 0
      )
    }
  }
}

dim(detection_array) # Should be num_sites x num_weeks x num_species

```

## Prepare Site Covariates
```{r SiteCov}
# Duplicate CameraLoc to be used in Objective 2
CameraLoc_O2 <- CameraLoc

# Standardize the covariates
CameraLoc <- CameraLoc %>%
  dplyr::select(-Plot, -Camera, -Lat, -Long, -Status, - Start_Date)
covariates_matrix <- as.matrix(CameraLoc)
rownames(covariates_matrix) <- site_names

# Standardizing covariates
covariates_matrix <- scale(covariates_matrix)

## Correlation
cor_mat <- cor(covariates_matrix)

# Create week matrix for covariate structure [site x week]
week_vals <- unique(observations_species$week)
week_matrix <- matrix(rep(week_vals, each = num_sites), nrow = num_sites, ncol = num_weeks, byrow = FALSE)

# Create detection covariate list
det.covs <- list(
  shrub_cover = covariates_matrix[, "total_shrub_cover"],
  veg_height = covariates_matrix[, "avg_veg_height"],
  week = week_matrix
)

# Remove dash and convert to numeric
week_numeric <- as.numeric(gsub("-", "", det.covs$week))

## Scale and center week_numeric
week_numeric <- scale(week_numeric)

# Reshape into the original 32x36 matrix
det.covs$week <- matrix(week_numeric, nrow = 32, ncol = 36)

det.covs$Auth <- matrix(CameraLoc$Auth, nrow = 32, ncol = 36)
str(det.covs)

```

## Combine data into a list
This requires combining the presence data and the site covariate data into a single list. This also means that the presence data is in a 3-d format.
```{r Combine_List_Data}
# Combine into a named list
data_list <- list(
  y = detection_array,
  occ.covs = covariates_matrix,
  det.covs = det.covs
  #,
  #coords = coords
)

str(data_list)
```

## Convert occupancy and detection covariates to a dataframe (from matrix)

```{r Convert_DF}
# Convert occupancy and detection covariates to a dataframe
data_list[["occ.covs"]] <- as.data.frame(data_list[["occ.covs"]])
data_list[["occ.covs"]]$total_shrub_cover <- as.numeric(data_list[["occ.covs"]]$total_shrub_cover)

#data_list[["det.covs"]] <- as.data.frame(data_list[["det.covs"]])
#data_list[["det.covs"]]$total_shrub_cover <- as.numeric(data_list[["det.covs"]]$total_shrub_cover)

# Make species the first dimension
data_list$y <- aperm(data_list$y, c(3, 1, 2))

dimnames(data_list$y) <- list(species = dimnames(data_list$y)[[1]],
                              site = dimnames(data_list$y)[[2]],
                              week = dimnames(data_list$y)[[3]])

str(data_list)
```

## Site Covariates
```{r Site_Covariates}
# Define detection formulas
det.null <- ~ 1
det.full <- ~ shrub_cover + veg_height + week + (1 | Auth)
det.cover <- ~ shrub_cover + veg_height + (1 | Auth)
det.week <- ~ week + (1 | Auth)
det.week.quad <- ~ week + I(week^2) + (1 | Auth)
det.full.quad <- ~ shrub_cover + veg_height + week + I(week^2) + (1 | Auth)


# Define occupancy formulas
occ.null <- ~ 1

occ.full <- ~ #Cogon_Patch_Size + 
  Veg_shannon_index + total_shrub_cover +
  Avg_Cogongrass_Cover + 
  #Tree_Density + 
  Avg_Canopy_Cover +
  avg_veg_height + CWD_presence + (1 | Auth)

occ.full.quad <- ~ #Cogon_Patch_Size + 
  Veg_shannon_index + total_shrub_cover +
  Avg_Cogongrass_Cover + I(Avg_Cogongrass_Cover^2) +
  #Tree_Density + 
  Avg_Canopy_Cover +
  avg_veg_height + CWD_presence + (1 | Auth)

occ.cover <- ~ Avg_Cogongrass_Cover + total_shrub_cover +
  avg_veg_height + CWD_presence + (1 | Auth)

occ.canopy <- ~ #Tree_Density + 
  Avg_Canopy_Cover + Avg_Cogongrass_Cover + (1 | Auth)

occ.canopy.quad <- ~ #Tree_Density + 
  Avg_Canopy_Cover + Avg_Cogongrass_Cover + I(Avg_Cogongrass_Cover^2) + (1 | Auth)

occ.move <- ~ #Cogon_Patch_Size + 
  Avg_Cogongrass_Cover + total_shrub_cover + (1 | Auth)

occ.move.quad <- ~ #Cogon_Patch_Size + 
  Avg_Cogongrass_Cover + I(Avg_Cogongrass_Cover^2) +
  total_shrub_cover + (1 | Auth)

occ.forage <- ~ Veg_shannon_index + Avg_Cogongrass_Cover + (1 | Auth)

occ.forage.quad <- ~ Veg_shannon_index + Avg_Cogongrass_Cover +
  I(Avg_Cogongrass_Cover^2) + (1 | Auth)

occ.cogon <- ~ Avg_Cogongrass_Cover + (1 | Auth)

occ.cogon.quad <- ~ Avg_Cogongrass_Cover + I(Avg_Cogongrass_Cover^2) + (1 | Auth)
```

## Priors
```{r Priors}
priors <- list(
  # Community-level effects
  beta.comm.normal  = list(mean = 0, var = 2.73),
  alpha.comm.normal = list(mean = 0, var = 2.73),

  # Species-level variance (hierarchical shrinkage)
  tau.sq.beta.ig  = list(2, 1),
  tau.sq.alpha.ig = list(2, 1)#,

  # Spatial variance
  #sigma.sq.ig = list(3, 2),

  # Spatial decay
  #phi.unif = list(0.05, 3) # 0.05 (5km) local & 3 (300 km) large-scale
)

```

## Run the occupancy model
```{r OccupancyModel, include=FALSE}
ms_null_null <- msPGOcc(
  occ.formula = occ.null,
  det.formula = det.null,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)


summary(ms_null_null)

# Includes all covariates of detection and occupancy
ms_full_full <- msPGOcc(
  occ.formula = occ.full,
  det.formula = det.full,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)


summary(ms_full_full)


#Includes cover for detection and only foraging for occupancy
ms_cover_forage <- msPGOcc(
  occ.formula = occ.forage,
  det.formula = det.cover,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_cover_forage)

# Includes movement covariates of occupancy and cover for detection
ms_cover_move <- msPGOcc(
  occ.formula = occ.move,
  det.formula = det.cover,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_cover_move)

#Includes cover covariate of detection and only canopy for occupancy
ms_cover_canopy <- msPGOcc(
  occ.formula = occ.canopy,
  det.formula = det.cover,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_cover_canopy)


#Includes cover covariate of detection and only quadratic canopy for occupancy
ms_cover_canopyQ <- msPGOcc(
  occ.formula = occ.canopy.quad,
  det.formula = det.cover,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_cover_canopyQ)



# Includes quadratic week covariate of detection and quadratic cogon for occupancy
ms_weekQ_cogonQ<- msPGOcc(
  occ.formula = occ.cogon.quad,
  det.formula = det.week.quad,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_weekQ_cogonQ)

# Includes quadratic week and full covariates of detection and all covariates and coverage for occupancy
ms_fullQ_cover<- msPGOcc(
  occ.formula = occ.cover,
  det.formula = det.full.quad,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_fullQ_cover)


#Includes quadratic week and full covariates of detection and only foraging for occupancy
ms_fullQ_forage<- msPGOcc(
  occ.formula = occ.forage,
  det.formula = det.full.quad,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_fullQ_forage)



#Includes quadratic week and full covariates of detection and quadratic of move for occupancy
ms_fullQ_moveQ<- msPGOcc(
  occ.formula = occ.move.quad,
  det.formula = det.full.quad,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_fullQ_moveQ)


# Includes quadratic week and full covariates of detection and all covariates and quadratic cogon for occupancy
ms_fullQ_fullQ<- msPGOcc(
  occ.formula = occ.full.quad,
  det.formula = det.full.quad,
  data = data_list,
  priors = priors,
  n.samples = 10000,
  n.burn = 3000,
  n.thin = 2,
  n.chains = 4,
  n.report = 1000
)

summary(ms_fullQ_fullQ)
```

## Model Comparison
```{r Model_Comparison}
waicOcc(ms_full_full, by.sp = FALSE)
waicOcc(ms_null_null, by.sp = FALSE)
waicOcc(ms_cover_forage, by.sp = FALSE)
waicOcc(ms_cover_move, by.sp = FALSE)
waicOcc(ms_cover_canopy, by.sp = FALSE)
waicOcc(ms_cover_canopyQ, by.sp = FALSE)
waicOcc(ms_weekQ_cogonQ, by.sp = FALSE)
waicOcc(ms_fullQ_moveQ, by.sp = FALSE)
waicOcc(ms_fullQ_cover, by.sp = FALSE)
waicOcc(ms_fullQ_forage, by.sp = FALSE)
waicOcc(ms_fullQ_fullQ, by.sp = FALSE)
```

## Model Validation
This test explains how well the model fits that data at the community and species level. I believe 0.5 is the target p-value, though how far from this number is considered adequate, I do not know yet. I believe this is a good place to check when thinking about which species we include in the model.
```{r Model_Validation}
ppc_ft <- ppcOcc(
  ms_cover_canopyQ,
  fit.stat = "freeman-tukey",
  group = 1
)

ppc_chisq <- ppcOcc(
  ms_cover_canopyQ,
  fit.stat = "chi-square",
  group = 1
)

summary(ppc_ft)
summary(ppc_chisq)

```

## Posterior Checks
```{r Posterior_Checks}
n.sp       <- dim(data_list$y)[1]
n.weeks    <- dim(data_list$y)[3]
sp.names   <- dimnames(data_list$y)[[1]]
plot.names <- dimnames(data_list$y)[[2]]

# Named lookup: scientific name -> common name
common.names <- c(
  "Canis_latrans"          = "Coyote",
  "Lynx_rufus"             = "Bobcat",
  "Procyon_lotor"          = "Raccoon",
  "Didelphis_virginiana"   = "Virginia Opossum",
  "Sylvilagus_floridanus"  = "Eastern Cottontail Rabbit",
  "Meleagris_gallopavo"    = "Wild Turkey",
  "Sciurus_carolinensis"   = "Grey Squirrel"
)

# ── Run PPC ───────────────────────────────────────────────────────────────
# group = 1: fit.y.group.quants = [5 x n.species x n.sites]
# group = 2: fit.y.group.quants = [5 x n.species x n.weeks]

ppc_ft_site <- ppcOcc(ms_cover_canopyQ, fit.stat = "freeman-tukey", group = 1)
ppc_ft_week <- ppcOcc(ms_cover_canopyQ, fit.stat = "freeman-tukey", group = 2)

ppc_chisq_site <- ppcOcc(ms_cover_canopyQ, fit.stat = "chi-square", group = 1)
ppc_chisq_week <- ppcOcc(ms_cover_canopyQ, fit.stat = "chi-square", group = 2)

summary(ppc_ft_site)
summary(ppc_ft_week)
summary(ppc_chisq_site)
summary(ppc_chisq_week)

# ── Scatter plot: replicated vs. observed fit statistic ───────────────────

make_scatter <- function(i) {
  ppc.df <- data.frame(
    fit     = ppc_ft_site$fit.y[ , i],
    fit.rep = ppc_ft_site$fit.y.rep[ , i]
  )
  ppc.df$color <- ifelse(ppc.df$fit.rep > ppc.df$fit, "lightsalmon", "lightskyblue1")

  ggplot(ppc.df, aes(x = fit, y = fit.rep)) +
    geom_point(aes(fill = color), shape = 21, color = "grey40", size = 1.8) +
    geom_abline(slope = 1, intercept = 0, color = "black") +
    scale_fill_identity() +
    labs(
      title = common.names[sp.names[i]],
      x = "True", y = "Replicate"
    ) +
    theme_bw(base_size = 10) +
    theme(plot.title = element_text(size = 9))
}

# ── site-level discrepancy ───────────────────────────────────────────
make_site_discrepancy <- function(i) {
  diff.fit <- ppc_ft_site$fit.y.rep.group.quants[3, i, ] -
              ppc_ft_site$fit.y.group.quants[3, i, ]

  df <- data.frame(
    x    = seq_along(diff.fit),
    diff = diff.fit,
    site = plot.names
  )

  ggplot(df, aes(x = x, y = diff)) +
    geom_point(shape = 19, size = 1.8) +
    geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
    scale_x_continuous(
      breaks = df$x,
      labels = df$site
    ) +
    labs(
      title = common.names[sp.names[i]],
      x = NULL, y = "Replicate - True Discrepancy"
    ) +
    theme_bw(base_size = 10) +
    theme(
      axis.text.x  = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 6),
      plot.title   = element_text(size = 9)
    )
}

# ── week-level discrepancy ───────────────────────────────────────────
make_week_discrepancy <- function(i) {
  diff.fit <- ppc_ft_week$fit.y.rep.group.quants[3, i, ] -
              ppc_ft_week$fit.y.group.quants[3, i, ]

  df <- data.frame(
    week = 1:n.weeks,
    diff = diff.fit
  )

  ggplot(df, aes(x = week, y = diff)) +
    geom_point(shape = 19, size = 1.8) +
    geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
    scale_x_continuous(breaks = 1:n.weeks) +
    labs(
      title = common.names[sp.names[i]],
      x = "Week", y = "Replicate - True Discrepancy"
    ) +
    theme_bw(base_size = 10) +
    theme(plot.title = element_text(size = 9))
}

# ── Assemble panels with patchwork ───────────────────────────────────────────
assemble_panel <- function(plot_fn, title_label) {
  plots <- lapply(seq_len(n.sp), plot_fn)
  wrap_plots(plots, ncol = ceiling(n.sp / 2)) +
    plot_annotation(title = title_label,
                    theme = theme(plot.title = element_text(size = 12, face = "bold")))
}

panel_scatter  <- assemble_panel(make_scatter,           "")
panel_site     <- assemble_panel(make_site_discrepancy,  "")
panel_week     <- assemble_panel(make_week_discrepancy,  "")

# ── Stack the three panels and save ──────────────────────────────────────────
ggsave(
  filename = "ppc_scatter.png",
  plot     = panel_scatter,
  width    = 12,
  height   = 9,
  dpi      = 300
)

ggsave(
  filename = "ppc_site_discrepancy.png",
  plot     = panel_site,
  width    = 12,
  height   = 9,
  dpi      = 300
)

ggsave(
  filename = "ppc_week_discrepancy.png",
  plot     = panel_week,
  width    = 12,
  height   = 9,
  dpi      = 300
)
```


## Proability of Direction- Occupancy
```{r Probability_Direction_occ}
pd <- function(x) {
  max(mean(x > 0), mean(x < 0))
}


pd_table <- apply(
  ms_cover_canopyQ$beta.samples,
  2,
  pd
)

head(pd_table)

beta_means <- colMeans(ms_cover_canopyQ$beta.samples)

beta_samps <- ms_cover_canopyQ$beta.samples

pd_vals <- apply(beta_samps, 2, pd)
beta_means <- colMeans(beta_samps)

# extract specific quantiles
get_quantile <- function(samples, prob) {
  apply(samples, 2, quantile, probs = prob)
}

results_species <- data.frame(
  parameter     = colnames(beta_samps),
  mean          = beta_means,
  pd            = pd_vals,
  lower_95_cri  = get_quantile(beta_samps, 0.025),
  upper_95_cri  = get_quantile(beta_samps, 0.975)
)

head(results_species)

beta_comm <- ms_cover_canopyQ$beta.comm.samples

results_comm <- data.frame(
  parameter     = colnames(beta_comm),
  mean          = colMeans(beta_comm),
  pd            = apply(beta_comm, 2, pd),
  lower_95_cri  = get_quantile(beta_comm, 0.025),
  upper_95_cri  = get_quantile(beta_comm, 0.975)
)

results_comm$pd_95 <- results_comm$pd >= 0.95


doc <- read_docx()


ft_species <- flextable(results_species) %>%
  colformat_double(digits = 4) %>%
  set_header_labels(
    parameter    = "Parameter",
    mean         = "Mean",
    pd           = "pd",
    lower_95_cri = "Lower 95% CrI",
    upper_95_cri = "Upper 95% CrI"
  ) %>%
  theme_booktabs() %>%
  autofit() %>%
  bold(i = ~ pd >= 0.95, bold = TRUE, part = "body")

doc <- doc %>%
  body_add_par("Species-Level Results", style = "heading 1") %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_species)


ft_comm <- flextable(results_comm) %>%
  colformat_double(digits = 4) %>%
  set_header_labels(
    parameter    = "Parameter",
    mean         = "Mean",
    pd           = "pd",
    lower_95_cri = "Lower 95% CrI",
    upper_95_cri = "Upper 95% CrI"
  ) %>%
  theme_booktabs() %>%
  autofit() %>%
  bold(i = ~ pd >= 0.95, bold = TRUE, part = "body")

doc <- doc %>%
  body_add_break() %>%
  body_add_par("Community-Level Results", style = "heading 1") %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_comm)


print(doc, target = "occ_model_results.docx")
```

## Probability of Direction- Detection
```{r Probability_Direction_det}
pd <- function(x) {
  max(mean(x > 0), mean(x < 0))
}

alpha_samps <- ms_cover_canopyQ$alpha.samples

pd_vals_alpha <- apply(alpha_samps, 2, pd)
alpha_means  <- colMeans(alpha_samps)

results_species_det <- data.frame(
  parameter = colnames(alpha_samps),
  mean = alpha_means,
  pd = pd_vals_alpha,
  pd_95 = pd_vals_alpha >= 0.95
)

head(results_species_det)

alpha_comm <- ms_cover_canopyQ$alpha.comm.samples

results_comm_det <- data.frame(
  parameter = colnames(alpha_comm),
  mean = colMeans(alpha_comm),
  pd = apply(alpha_comm, 2, pd)
)

results_comm_det$pd_95 <- results_comm_det$pd >= 0.95

results_comm_det

```

## Plot Species Occupancy
```{r Plot_Species_Occupancy, eval=FALSE, include=FALSE}
summary(ms_cover_canopyQ) # Summary of parameter estimates
names(ms_cover_canopyQ)
str(ms_cover_canopyQ$beta.samples)

mean(ms_cover_canopyQ$beta.samples[, 5] > 0)

MCMCplot(ms_cover_canopyQ$beta.samples, ref_ovl = TRUE, ci = c(50, 95)) # Occupancy
MCMCplot(ms_cover_canopyQ$alpha.samples, ref_ovl = TRUE, ci = c(50, 95)) # Detection

# unscaled cogongrass values
cogon.pred.vals <- seq(min(data_list$occ.covs$Avg_Cogongrass_Cover),
                       max(data_list$occ.covs$Avg_Cogongrass_Cover),
                       length.out = 100)

# Scale using the same transformation used in the model
cogon.mean <- mean(data_list$occ.covs$Avg_Cogongrass_Cover)
cogon.sd   <- sd(data_list$occ.covs$Avg_Cogongrass_Cover)

cogon.pred.scale <- (cogon.pred.vals - cogon.mean) / cogon.sd

# quadratic term (scaled^2)
cogon.pred.scale2 <- cogon.pred.scale^2

# Prediction matrix
pred.df <- as.matrix(data.frame(
  intercept = 1,
  Avg_Cogongrass_Cover = cogon.pred.scale,
  I.Avg_Cogongrass_Cover.2. = cogon.pred.scale2,
 # CWD_presence = 0,
 # Cogon_Patch_Size = 0,
#  Veg_shannon_index = 0,
#  total_shrub_cover = 0,
#  Tree_Density = 0,
  Avg_Canopy_Cover = 0
#  avg_veg_height = 0,
#  Auth = 0
))


# Predict occupancy from the model
out.pred <- predict(ms_cover_canopyQ,
                    X.0 = pred.df,
                    ignore.RE = TRUE)


# Convert posterior samples to quantiles
psi.0.quants <- apply(out.pred$psi.0.samples, 3,
                      function(x) quantile(x, prob = c(0.025, 0.5, 0.975)))

# Format for plotting
psi.plot.dat <- data.frame(
  psi.med = psi.0.quants[2, ],
  psi.low = psi.0.quants[1, ],
  psi.high = psi.0.quants[3, ],
  Avg_Cogongrass_Cover = cogon.pred.vals
)

# Plot
ggplot(psi.plot.dat, aes(x = Avg_Cogongrass_Cover, y = psi.med)) +
  geom_ribbon(aes(ymin = psi.low, ymax = psi.high), fill = "grey70") +
  geom_line(size = 1.1) +
  theme_bw() +
  labs(x = "Average Cogongrass Cover (%)", y = "Predicted Occupancy") +
  scale_y_continuous(limits = c(0,1))


```



## Batched Parameter Estimates
```{r Batched_Parameter_Estimates}
summary(ms_cover_canopyQ) # Summary of parameter estimates
names(ms_cover_canopyQ)
str(ms_cover_canopyQ$beta.samples)

mean(ms_cover_canopyQ$beta.samples[, 5] > 0)

MCMCplot(ms_cover_canopyQ$beta.samples, ref_ovl = TRUE, ci = c(50, 95)) # Occupancy
MCMCplot(ms_cover_canopyQ$alpha.samples, ref_ovl = TRUE, ci = c(50, 95)) # Detection

## Occupancy
# Total number of parameters
n_params <- ncol(ms_cover_canopyQ$beta.samples)

# Number of parameters to plot at a time
chunk_size <- 10

# Split and plot
for (i in seq(1, n_params, by = chunk_size)) {
  end <- min(i + chunk_size - 1, n_params)
  param_names <- colnames(ms_cover_canopyQ$beta.samples)[i:end]
  
  # Set filename
  file_name <- paste0("MCMCplot_Occupancy_Params_", i, "_to_", end, ".png")
  
  # Save plot to PNG
  png(filename = file_name, width = 1200, height = 800, res = 150)
  
  MCMCplot(ms_cover_canopyQ$beta.samples[, param_names],
           ref_ovl = TRUE,
           ci = c(50, 95),
           main = paste0("Occupancy Parameters: ", i, " to ", end))
  
  dev.off()
}


## Detection 
# Number of parameters
n_params <- ncol(ms_cover_canopyQ$alpha.samples)

# Number of parameters to plot at a time
chunk_size <- 10

# Split and plot
for (i in seq(1, n_params, by = chunk_size)) {
  end <- min(i + chunk_size - 1, n_params)
  param_names <- colnames(ms_cover_canopyQ$alpha.samples)[i:end]
  
  # Set filename
  file_name <- paste0("MCMCplot_Detection_Params_", i, "_to_", end, ".png")
  
  # Save plot to PNG
  png(filename = file_name, width = 1200, height = 800, res = 150)
  
  MCMCplot(ms_cover_canopyQ$alpha.samples[, param_names, drop = FALSE],
         ref_ovl = TRUE,
         ci = c(50, 95),
         main = paste0("Detection Parameters: ", i, " to ", end))
  
  dev.off()
}
```

## Facet Plots
```{r Facet_Plots}
plot_mcmc_faceted_groups <- function(samples_matrix,
                                     chunk_size = 8,
                                     title_text = "Parameters",
                                     pd_cutoff = 0.95) {
  
  param_names <- colnames(samples_matrix)
  
  
  df <- as.data.frame(samples_matrix)
  df$Iteration <- 1:nrow(df)
  
  df_long <- df %>%
    pivot_longer(-Iteration,
                 names_to = "Parameter",
                 values_to = "Value") %>%
    mutate(
      ParamIndex = match(Parameter, param_names),
      Group = ceiling(ParamIndex / chunk_size)
    )
  
  # Posterior summaries + probability of direction
  summary_df <- df_long %>%
    group_by(Parameter, Group) %>%
    summarise(
      mean = mean(Value),
      l95 = quantile(Value, 0.025),
      u95 = quantile(Value, 0.975),
      l50 = quantile(Value, 0.25),
      u50 = quantile(Value, 0.75),
      pd = max(mean(Value > 0), mean(Value < 0)),
      .groups = "drop"
    ) %>%
    mutate(
      Credible = pd >= pd_cutoff
    )
  
  
  summary_df <- summary_df %>%
    group_by(Group) %>%
    mutate(Parameter = factor(Parameter,
                               levels = rev(unique(Parameter)))) %>%
    ungroup()
  
  
  ggplot(summary_df,
         aes(x = mean, y = Parameter)) +
    
    geom_vline(xintercept = 0, linetype = "dashed") +
    
    # 95% CI
    geom_errorbarh(aes(xmin = l95,
                       xmax = u95,
                       size = Credible),
                   height = 0) +
    
    # 50% CI
    geom_errorbarh(aes(xmin = l50,
                       xmax = u50,
                       size = Credible),
                   height = 0) +
    
    # Posterior mean
    geom_point(aes(fill = Credible),
               shape = 21,
               size = 3,
               stroke = 1) +
    
    facet_wrap(~Group, scales = "free_y", ncol = 2) +
    
    scale_size_manual(values = c("TRUE" = 1.2,
                                 "FALSE" = 0.5),
                      guide = "none") +
    
    scale_fill_manual(values = c("TRUE" = "black",
                                 "FALSE" = "white"),
                      guide = "none") +
    
    labs(x = "Parameter Estimate",
         y = "",
         title = title_text) +
    
    theme_bw() +
    theme(
      strip.background = element_blank(),
      strip.text = element_blank(), 
      axis.text.y = element_text(size = 10)
    )
}

plot_mcmc_faceted_groups(
  ms_cover_canopyQ$beta.samples,
  chunk_size = 8,
  title_text = "",
  pd_cutoff = 0.95
)

ggsave("mcmc_faceted_groups_occupancy.png",
       width = 10, height = 8, dpi = 300)

plot_mcmc_faceted_groups(
  ms_cover_canopyQ$alpha.samples,
  chunk_size = 8,
  title_text = "",
  pd_cutoff = 0.95
)

ggsave("mcmc_faceted_groups_detection.png",
       width = 10, height = 8, dpi = 300)

```


# Plot Species level relationship with cogongrass cover
```{r Species_Level_Cogon}
# Collecting scaled atrributes
cogon.mean <- attr(covariates_matrix, "scaled:center")["Avg_Cogongrass_Cover"]
cogon.sd <- attr(covariates_matrix, "scaled:scale")["Avg_Cogongrass_Cover"]

cogon.mean <- mean(CameraLoc$Avg_Cogongrass_Cover)
cogon.sd <- sd(CameraLoc$Avg_Cogongrass_Cover)

# Unscaled prediction values
cogon.pred.vals <- seq(0, 100, length.out = 100)

# Scale prediction values
cogon.pred.scale <- (cogon.pred.vals - cogon.mean) / cogon.sd
cogon.pred.scale2 <- cogon.pred.scale^2

canopy.mean <- attr(covariates_matrix, "scaled:center")["Avg_Canopy_Cover"]
canopy.sd <- attr(covariates_matrix, "scaled:scale")["Avg_Canopy_Cover"]

pred.df <- as.matrix(data.frame(
  intercept = 1,
  Avg_Cogongrass_Cover = cogon.pred.scale,
  I.Avg_Cogongrass_Cover.2. = cogon.pred.scale2,
  Avg_Canopy_Cover = 0  
))

out.pred <- predict(
  ms_cover_canopyQ,
  X.0 = pred.df,
  ignore.RE = TRUE
)


# Extract the species-specific betas for Coyote
coyote_index <- which(ms_cover_canopyQ$sp.names == "Canis_latrans")

# Get column indices for the parameters we need
intercept_col <- paste0("(Intercept)-", "Canis_latrans")
cogon_col <- paste0("Avg_Cogongrass_Cover-", "Canis_latrans")
cogon2_col <- paste0("I(Avg_Cogongrass_Cover^2)-", "Canis_latrans")
canopy_col <- paste0("Avg_Canopy_Cover-", "Canis_latrans")

# Extract posterior samples for each coefficient
beta_intercept <- ms_cover_canopyQ$beta.samples[, intercept_col]
beta_cogon <- ms_cover_canopyQ$beta.samples[, cogon_col]
beta_cogon2 <- ms_cover_canopyQ$beta.samples[, cogon2_col]
beta_canopy <- ms_cover_canopyQ$beta.samples[, canopy_col]

# Create prediction gradient (0 to 100%)
cogon.pred.vals <- seq(0, 100, length.out = 1000)

# Scale using same transformation
cogon.pred.scale <- (cogon.pred.vals - cogon.mean) / cogon.sd
cogon.pred.scale2 <- cogon.pred.scale^2

n_samples <- nrow(ms_cover_canopyQ$beta.samples)
n_pred <- length(cogon.pred.vals)

logit_psi <- matrix(NA, nrow = n_samples, ncol = n_pred)

for(i in 1:n_samples) {
  logit_psi[i, ] <- beta_intercept[i] + 
                    beta_cogon[i] * cogon.pred.scale + 
                    beta_cogon2[i] * cogon.pred.scale2 + 
                    beta_canopy[i] * 0  # Setting to mean (0 on scaled)
}

# Convert to probability scale
psi_samples <- plogis(logit_psi)

# Calculate quantiles across MCMC samples for each prediction point
psi.coyote.quants <- apply(psi_samples, 2, quantile, 
                            probs = c(0.025, 0.5, 0.975))

psi.coyote.dat <- data.frame(
  Avg_Cogongrass_Cover = cogon.pred.vals,
  psi.low  = psi.coyote.quants[1, ],
  psi.med  = psi.coyote.quants[2, ],
  psi.high = psi.coyote.quants[3, ]
)

ggplot(psi.coyote.dat, aes(x = Avg_Cogongrass_Cover, y = psi.med)) +
  geom_ribbon(aes(ymin = psi.low, ymax = psi.high), fill = "grey70", alpha = 0.5) +
  geom_line(size = 1.2) +
  theme_bw() +
  labs(x = "Average Cogongrass Cover (%)",
       y = "Occupancy Probability (Canis latrans)") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 100))

summary(psi.coyote.dat)
```


## Community Occupancy
```{r Plot_Community_Occupancy}
# Community-level (average across species)
comm_intercept <- ms_cover_canopyQ$beta.comm.samples[, "(Intercept)"]
comm_cogon <- ms_cover_canopyQ$beta.comm.samples[, "Avg_Cogongrass_Cover"]
comm_cogon2 <- ms_cover_canopyQ$beta.comm.samples[, "I(Avg_Cogongrass_Cover^2)"]
comm_canopy <- ms_cover_canopyQ$beta.comm.samples[, "Avg_Canopy_Cover"]

# Same prediction approach
logit_psi_comm <- matrix(NA, nrow = nrow(ms_cover_canopyQ$beta.comm.samples), 
                         ncol = n_pred)

for(i in 1:nrow(ms_cover_canopyQ$beta.comm.samples)) {
  logit_psi_comm[i, ] <- comm_intercept[i] + 
                         comm_cogon[i] * cogon.pred.scale + 
                         comm_cogon2[i] * cogon.pred.scale2 + 
                         comm_canopy[i] * 0
}

psi_comm_samples <- plogis(logit_psi_comm)

psi.comm.quants <- apply(psi_comm_samples, 2, quantile, 
                         probs = c(0.025, 0.5, 0.975))

psi.comm.dat <- data.frame(
  Avg_Cogongrass_Cover = cogon.pred.vals,
  psi.low  = psi.comm.quants[1, ],
  psi.med  = psi.comm.quants[2, ],
  psi.high = psi.comm.quants[3, ]
)

ggplot(psi.comm.dat, aes(x = Avg_Cogongrass_Cover, y = psi.med)) +
  geom_ribbon(aes(ymin = psi.low, ymax = psi.high), fill = "grey70", alpha = 0.5) +
  geom_line(size = 1.2) +
  theme_bw() +
  labs(x = "Average Cogongrass Cover (%)",
       y = "Community Occupancy Probability") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 100))


```


## Loop through all species in the community
```{r Loop_Species_Occupancy}
# Prediction gradient
cogon.pred.vals <- seq(0, 100, length.out = 1000)
cogon.pred.scale <- (cogon.pred.vals - cogon.mean) / cogon.sd
cogon.pred.scale2 <- cogon.pred.scale^2

n_samples <- nrow(ms_cover_canopyQ$beta.samples)
n_pred <- length(cogon.pred.vals)

# List to store results
species_predictions <- list()
species_plots <- list()

# Loop through all species
for(sp in ms_cover_canopyQ$sp.names) {
  
  # Get column names for this species
  intercept_col <- paste0("(Intercept)-", sp)
  cogon_col <- paste0("Avg_Cogongrass_Cover-", sp)
  cogon2_col <- paste0("I(Avg_Cogongrass_Cover^2)-", sp)
  canopy_col <- paste0("Avg_Canopy_Cover-", sp)
  
  # Extract posterior samples
  beta_intercept <- ms_cover_canopyQ$beta.samples[, intercept_col]
  beta_cogon <- ms_cover_canopyQ$beta.samples[, cogon_col]
  beta_cogon2 <- ms_cover_canopyQ$beta.samples[, cogon2_col]
  beta_canopy <- ms_cover_canopyQ$beta.samples[, canopy_col]
  
  # Calculate linear predictor
  logit_psi <- matrix(NA, nrow = n_samples, ncol = n_pred)
  
  for(i in 1:n_samples) {
    logit_psi[i, ] <- beta_intercept[i] + 
                      beta_cogon[i] * cogon.pred.scale + 
                      beta_cogon2[i] * cogon.pred.scale2 + 
                      beta_canopy[i] * 0
  }
  
  # Convert to probability
  psi_samples <- plogis(logit_psi)
  
  # Calculate quantiles
  psi.quants <- apply(psi_samples, 2, quantile, probs = c(0.025, 0.5, 0.975))
  
  # Store results
  species_predictions[[sp]] <- data.frame(
    species = sp,
    Avg_Cogongrass_Cover = cogon.pred.vals,
    psi.low  = psi.quants[1, ],
    psi.med  = psi.quants[2, ],
    psi.high = psi.quants[3, ]
  )
  
  # Create plot
  species_plots[[sp]] <- ggplot(species_predictions[[sp]], 
                                 aes(x = Avg_Cogongrass_Cover, y = psi.med)) +
    geom_ribbon(aes(ymin = psi.low, ymax = psi.high), fill = "grey70", alpha = 0.5) +
    geom_line(size = 1.2) +
    theme_bw() +
    labs(x = "Average Cogongrass Cover (%)",
         y = "Occupancy Probability",
         title = gsub("_", " ", sp)) +  # Replace underscores with spaces
    scale_y_continuous(limits = c(0, 1)) +
    scale_x_continuous(limits = c(0, 100))
}

# View individual plots
species_plots[["Canis_latrans"]]
species_plots[["Procyon_lotor"]]

# Combine all plots into a grid
all_species_plot <- wrap_plots(species_plots, ncol = 4)
all_species_plot

# Save combined plot
ggsave("all_species_cogongrass_response.png", 
       all_species_plot, 
       width = 12, height = 16, dpi = 300)

# Combine all predictions into one dataframe for faceted plot
all_predictions <- do.call(rbind, species_predictions)

# Create faceted plot
faceted_plot <- ggplot(all_predictions, aes(x = Avg_Cogongrass_Cover, y = psi.med)) +
  geom_ribbon(aes(ymin = psi.low, ymax = psi.high), fill = "grey70", alpha = 0.5) +
  geom_line(size = 1.2) +
  facet_wrap(~ species, ncol = 2, labeller = labeller(species = function(x) gsub("_", " ", x))) +
  theme_bw() +
  labs(x = "Average Cogongrass Cover (%)",
       y = "Occupancy Probability") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 100)) +
  theme(strip.text = element_text(face = "italic"))  # Italicize species names

faceted_plot

ggsave("all_species_faceted.png", 
       faceted_plot, 
       width = 10, height = 12, dpi = 300)
```

## Species Lookup
```{r Species_Lookup}
SpeciesLookup <- tibble::tribble(
  ~Name, ~Scientific_Name, ~Common_Name,
  "Canis_latrans", "Canis latrans", "Coyote",
  "Didelphis_virginiana", "Didelphis virginiana", "Virginia Opossum",
  "Procyon_lotor", "Procyon lotor", "Raccoon",
  "Lynx_rufus", "Lynx rufus", "Bobcat",
  "Meleagris_gallopavo", "Meleagris gallopavo", "Wild Turkey",
  "Sciurus_carolinensis", "Sciurus carolinensis", "Eastern Gray Squirrel",
  "Sylvilagus_floridanus", "Sylvilagus floridanus", "Eastern Cottontail Rabbit",
)

```

## Varying cogongrass and canopy cover
```{r Varying_Cogon_Canopy}
# Prediction gradient for cogongrass
cogon.pred.vals <- seq(0, 100, length.out = 1000)
cogon.pred.scale <- (cogon.pred.vals - cogon.mean) / cogon.sd
cogon.pred.scale2 <- cogon.pred.scale^2

# Define canopy cover scenarios with factor levels
canopy_scenarios <- data.frame(
  scenario = factor(c("Low", "Average", "High"),
                   levels = c("Low", "Average", "High")),
  canopy_scaled = c(-1, 0, 1),
  color = c("#D55E00", "#009E73", "#0072B2")
)

n_samples <- nrow(ms_cover_canopyQ$beta.samples)
n_pred <- length(cogon.pred.vals)

# Lists for results
species_predictions <- list()
species_plots <- list()

# Loop through all species
for(sp in ms_cover_canopyQ$sp.names) {
  
  # Get column names for this species
  intercept_col <- paste0("(Intercept)-", sp)
  cogon_col <- paste0("Avg_Cogongrass_Cover-", sp)
  cogon2_col <- paste0("I(Avg_Cogongrass_Cover^2)-", sp)
  canopy_col <- paste0("Avg_Canopy_Cover-", sp)
  
  # Extract posterior samples
  beta_intercept <- ms_cover_canopyQ$beta.samples[, intercept_col]
  beta_cogon <- ms_cover_canopyQ$beta.samples[, cogon_col]
  beta_cogon2 <- ms_cover_canopyQ$beta.samples[, cogon2_col]
  beta_canopy <- ms_cover_canopyQ$beta.samples[, canopy_col]
  
  # Store predictions for each canopy scenario
  scenario_predictions <- list()
  
  for(j in 1:nrow(canopy_scenarios)) {
    
    canopy_val <- canopy_scenarios$canopy_scaled[j]
    
    # Calculate linear predictor
    logit_psi <- matrix(NA, nrow = n_samples, ncol = n_pred)
    
    for(i in 1:n_samples) {
      logit_psi[i, ] <- beta_intercept[i] + 
                        beta_cogon[i] * cogon.pred.scale + 
                        beta_cogon2[i] * cogon.pred.scale2 + 
                        beta_canopy[i] * canopy_val
    }
    
    # Convert to probability
    psi_samples <- plogis(logit_psi)
    
    # Calculate quantiles
    psi.quants <- apply(psi_samples, 2, quantile, probs = c(0.025, 0.5, 0.975))
    
    # Store results
    scenario_predictions[[j]] <- data.frame(
      species = sp,
      scenario = canopy_scenarios$scenario[j],
      Avg_Cogongrass_Cover = cogon.pred.vals,
      psi.low  = psi.quants[1, ],
      psi.med  = psi.quants[2, ],
      psi.high = psi.quants[3, ]
    )
  }
  
  # Combine all scenarios
  species_predictions[[sp]] <- do.call(rbind, scenario_predictions)
  
  all_predictions <- all_predictions %>%
  left_join(SpeciesLookup, by = c("species" = "Name"))
  # Create plot
  species_plots[[sp]] <- ggplot(species_predictions[[sp]], 
                                 aes(x = Avg_Cogongrass_Cover, y = psi.med, 
                                     color = scenario, fill = scenario)) +
    geom_ribbon(aes(ymin = psi.low, ymax = psi.high), alpha = 0.2, color = NA) +
    geom_line(size = 1.2) +
    scale_color_manual(values = canopy_scenarios$color) +
    scale_fill_manual(values = canopy_scenarios$color) +
    theme_bw() +
    labs(x = "Average Cogongrass Cover (%)",
         y = "Occupancy Probability",
         title = gsub("_", " ", sp),
         color = "Canopy Cover",
         fill = "Canopy Cover") +
    scale_y_continuous(limits = c(0, 1)) +
    scale_x_continuous(limits = c(0, 100)) +
    theme(
  legend.position = "bottom",
  axis.text = element_text(size = 14),
  axis.title = element_text(size = 16),
  plot.title = element_text(size = 16),
  legend.text = element_text(size = 13),
  legend.title = element_text(size = 14)
)
  
}

# View individual plots
species_plots[["Canis_latrans"]]
species_plots[["Procyon_lotor"]]

# Combine all plots into a grid
all_species_plot <- wrap_plots(species_plots, ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom")

all_species_plot


ggsave("all_species_cogongrass_canopy.png", 
       all_species_plot, 
       width = 14, height = 16, dpi = 300)

# Combine all predictions into one dataframe
all_predictions <- do.call(rbind, species_predictions)

all_predictions <- all_predictions %>%
  left_join(
    SpeciesLookup %>% dplyr::select(Name, Common_Name),
    by = c("species" = "Name")
  )

all_predictions$scenario <- factor(all_predictions$scenario,
                                   levels = c("Low", "Average", "High"))

# Create faceted plot
faceted_plot <- ggplot(all_predictions, 
                       aes(x = Avg_Cogongrass_Cover, y = psi.med, 
                           color = scenario, fill = scenario)) +
  geom_ribbon(aes(ymin = psi.low, ymax = psi.high), alpha = 0.2, color = NA) +
  geom_line(size = 1.2) +
  facet_wrap(~ Common_Name, ncol = 2) +
  scale_color_manual(values = canopy_scenarios$color) +
  scale_fill_manual(values = canopy_scenarios$color) +
  theme_bw() +
  labs(x = "Average Cogongrass Cover (%)",
       y = "Occupancy Probability",
       color = "Canopy Cover",
       fill = "Canopy Cover") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 100)) +
  theme(
  strip.text = element_text(face = "bold", size = 14),
  legend.position = "bottom",
  axis.text = element_text(size = 14),
  axis.title = element_text(size = 16),
  legend.text = element_text(size = 13),
  legend.title = element_text(size = 14)
)


faceted_plot


ggsave("all_species_faceted_canopy.png", 
       faceted_plot, 
       width = 12, height = 14, dpi = 600)

```

## Community Occupancy Plot- Cogongrass and Canopy Cover
```{r Community_Cogon_Canopy}
# Prediction gradient for cogongrass
cogon.pred.vals <- seq(0, 100, length.out = 1000)
cogon.pred.scale <- (cogon.pred.vals - cogon.mean) / cogon.sd
cogon.pred.scale2 <- cogon.pred.scale^2

# Define canopy cover scenarios
canopy_scenarios <- data.frame(
  scenario = c("Low", "Average", "High"),
  canopy_scaled = c(-1, 0, 1),
  color = c("#D55E00", "#009E73", "#0072B2")
)

n_pred <- length(cogon.pred.vals)

# Get community-level parameters
comm_intercept <- ms_cover_canopyQ$beta.comm.samples[, "(Intercept)"]
comm_cogon <- ms_cover_canopyQ$beta.comm.samples[, "Avg_Cogongrass_Cover"]
comm_cogon2 <- ms_cover_canopyQ$beta.comm.samples[, "I(Avg_Cogongrass_Cover^2)"]
comm_canopy <- ms_cover_canopyQ$beta.comm.samples[, "Avg_Canopy_Cover"]

# Store predictions for each canopy scenario
community_predictions <- list()

for(j in 1:nrow(canopy_scenarios)) {
  
  canopy_val <- canopy_scenarios$canopy_scaled[j]
  
  # Calculate linear predictor
  logit_psi_comm <- matrix(NA, 
                           nrow = nrow(ms_cover_canopyQ$beta.comm.samples), 
                           ncol = n_pred)
  
  for(i in 1:nrow(ms_cover_canopyQ$beta.comm.samples)) {
    logit_psi_comm[i, ] <- comm_intercept[i] + 
                           comm_cogon[i] * cogon.pred.scale + 
                           comm_cogon2[i] * cogon.pred.scale2 + 
                           comm_canopy[i] * canopy_val
  }
  
  # Convert to probability
  psi_comm_samples <- plogis(logit_psi_comm)
  
  # Calculate quantiles
  psi.comm.quants <- apply(psi_comm_samples, 2, quantile, 
                           probs = c(0.025, 0.5, 0.975))
  
  # Store results
  community_predictions[[j]] <- data.frame(
    scenario = canopy_scenarios$scenario[j],
    Avg_Cogongrass_Cover = cogon.pred.vals,
    psi.low  = psi.comm.quants[1, ],
    psi.med  = psi.comm.quants[2, ],
    psi.high = psi.comm.quants[3, ]
  )
}

# Combine all scenarios
psi.comm.dat <- do.call(rbind, community_predictions)

# Reorder the scenario factor levels
psi.comm.dat$scenario <- factor(psi.comm.dat$scenario, 
                                levels = c("Low", "Average", "High"))

community_plot <- ggplot(psi.comm.dat, 
                         aes(x = Avg_Cogongrass_Cover, y = psi.med, 
                             color = scenario, fill = scenario)) +
  geom_ribbon(aes(ymin = psi.low, ymax = psi.high), alpha = 0.2, color = NA) +
  geom_line(size = 1.5) +
  scale_color_manual(values = canopy_scenarios$color) +
  scale_fill_manual(values = canopy_scenarios$color) +
  theme_bw(base_size = 14) +
  labs(x = "Average Cogongrass Cover (%)",
       y = "Community Occupancy Probability",
       color = "Canopy Cover",
       fill = "Canopy Cover") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 100)) +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))

community_plot

ggsave("community_cogongrass_canopy.png", 
       community_plot, 
       width = 10, height = 7, dpi = 300)

write.csv(psi.comm.dat, "community_predictions_by_canopy.csv", row.names = FALSE)
```

## Shared Plot
```{r Shared_Plot}
psi.comm.dat$species <- "Community"
psi.comm.dat$Common_Name <- "Community (All Species)"

all_predictions2 <- bind_rows(all_predictions, psi.comm.dat)

facet_labels <- all_predictions2 %>%
  distinct(Common_Name) %>%
  arrange(Common_Name) %>%
  mutate(label = LETTERS[1:n()])

all_predictions2 <- all_predictions2 %>%
  left_join(facet_labels, by = "Common_Name")

faceted_plot <- ggplot(all_predictions2, 
                       aes(x = Avg_Cogongrass_Cover, 
                           y = psi.med, 
                           color = scenario, 
                           fill = scenario)) +
  geom_ribbon(aes(ymin = psi.low, ymax = psi.high), alpha = 0.2, color = NA) +
  geom_line(size = 1.2) +
  
  geom_text(
    data = facet_labels,
    aes(x = 5, y = 0.95, label = label),
    inherit.aes = FALSE,
    fontface = "bold",
    size = 6
  ) +
  
  facet_wrap(~ Common_Name, ncol = 2) +
  scale_color_manual(values = canopy_scenarios$color) +
  scale_fill_manual(values = canopy_scenarios$color) +
  labs(
    x = "Average Cogongrass Cover (%)",
    y = "Occupancy Probability",
    color = "Canopy Cover",
    fill = "Canopy Cover"
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 100)) +
  theme_bw() +
  theme(
    strip.text = element_text(face = "bold", size = 14),
    legend.position = "bottom",
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )

print(faceted_plot)

ggsave("all_species_faceted_with_community.png", 
       faceted_plot, 
       width = 12, height = 14, dpi = 600)

```